Arrows for Parallel Computations

### 2017-11-25 paper/content/benchmarks.tex
.

Apart from Eden, all benchmarks and libaries were compiled with Stack's \footnote{see https://www.haskellstack.org/} lts-7.1 GHC compiler which is equivalent to a standard GHC 8.0.1 with the base package in version 4.9.0.0. Stack itself was used in version 1.3.2. For GpH in its Multicore variant we used the the parallel cabal package in version 3.2.1.0, while for the |Par| monad we used monad-par in version 0.3.4.8. For all Eden tests we used its GHC-Eden compiler in version 7.8.2 together with OpenMPI 1.6.5.

@@

\subsection{Test programs}

@@

\begin{table}
\centering
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{0.5em}
\begin{tabular}{lccll}
\toprule
Name & Area & Type & Origin & Citation \\
\midrule
\rmtest & Mathematics & \ensuremath{\Varid{parMap}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\jacobitest & Mathematics & \ensuremath{\Varid{workpool}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\torustest & Mathematics & \ensuremath{\Varid{torus}} & Eden & \citet{Eden:SkeletonBookChapter02}\\
\sudokutest & Puzzle & \ensuremath{\Varid{parMap}} & \ensuremath{\Conid{Par}} Monad & \citet{par-monad} 
\tablefootnote{actual code from: http://community.haskell.org/\~simonmar/par-tutorial.pdf and https://github.com/simonmar/parconc-examples}\\
\bottomrule
\end{tabular}
\label{tab:benches}
\caption{The benchmarks we use in this paper.}
\end{table}

@@

\enquote{Gentleman} is a standard Eden test program, developed
for their \ensuremath{\Varid{torus}} skeleton. It implements a parallel matrix
multiplication \citep{Gentleman1978}. We ported an Eden based version \citep{Eden:SkeletonBookChapter02} to PArrows.

A~parallel Sudoku solver was used by \citet{par-monad} to compare \ensuremath{\Conid{Par}} Monad
to GpH Haskell. We ported it to PArrows.

@@

The \ensuremath{\Conid{Par}} monad and GpH Haskell -- in its multicore version \cite{Marlow2009} --  can be executed on a shared
memory machines only. Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

@@

The PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup (Fig.~\ref{fig:bench-rm-sm}). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65, the
efficiency was 85.3\%. However, if we increase the number of
requested cores to be 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup does not increase that well. It is merely 15.99
for 32 tasks with PArrows/Eden. It is worse for other backends.  As
for 64 tasks, we obtain the speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. Efficiency is 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
cluster fails to execute two parallel Haskell processes with full-fledged
parallelism. For this reason, we did not regard hyper-threaded cores in
our speedup plots in Figs.~\ref{fig:bench-rm-sm}--\ref{fig:sudokuSMBenchmark}.

>>>>

Apart from Eden, all benchmarks and libaries were compiled with Stack's \footnote{see \url{https://www.haskellstack.org/}} lts-7.1 GHC compiler which is equivalent to a standard GHC 8.0.1 with the base package in version 4.9.0.0. Stack itself was used in version 1.3.2. For GpH in its Multicore variant we used the the parallel cabal package in version 3.2.1.0, while for the |Par| monad we used monad-par in version 0.3.4.8. For all Eden tests, we used its GHC-Eden compiler in version 7.8.2 together with OpenMPI 1.6.5.

Furtermore, all benchmarks were done with help of the bench \footnote{see \url{https://hackage.haskell.org/package/bench}} tool in version 1.0.2 which uses criterion (>=1.1.1.0 \&\& < 1.2) internally. All runtime data (mean runtime, max stddev, etc.) was collected with this tool if not mentioned otherwise.

@@

\subsection{Benchmarks}

@@

\begin{table}
\centering
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{0.5em}
\begin{tabular}{lccll}
\toprule
Name & Area & Type & Origin & Source \\
\midrule
\rmtest & Mathematics & \ensuremath{\Varid{parMap}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\jacobitest & Mathematics & \ensuremath{\Varid{workpool}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\torustest & Mathematics & \ensuremath{\Varid{torus}} & Eden & \citet{Eden:SkeletonBookChapter02}\\
\sudokutest & Puzzle & \ensuremath{\Varid{parMap}} & \ensuremath{\Conid{Par}} Monad & \citet{par-monad} 
\tablefootnote{actual code from: http://community.haskell.org/\~simonmar/par-tutorial.pdf and https://github.com/simonmar/parconc-examples}\\
\bottomrule
\end{tabular}
\label{tab:benches}
\caption{The benchmarks we use in this paper.}
\end{table}

@@

\enquote{Gentleman} is a standard Eden test program, developed
for their \ensuremath{\Varid{torus}} skeleton. It implements a Gentleman's algorithm for parallel matrix
multiplication \citep{Gentleman1978}. We ported an Eden based version \citep{Eden:SkeletonBookChapter02} to PArrows.

A~parallel Sudoku solver was used by \citet{par-monad} to compare \ensuremath{\Conid{Par}} Monad
to GpH, and we ported it to PArrows.

@@

The \ensuremath{\Conid{Par}} monad and GpH -- in its multicore version \cite{Marlow2009} --  can be executed on a shared
memory machines only. Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

@@

In preliminary tests, the PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup on up to 16 shared-memory cores (Fig.~\ref{fig:bench-rm-sm}). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65 giving a parallel
efficiency of 85.3\%. However, if we increased the number of
requested cores to 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup did not increase that well. It was merely 15.99
for 32 tasks with PArrows/Eden. This was worse for other backends.  As
for 64 tasks, we obtained a speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. The parallel eficiencies were 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
cluster fails to execute two parallel Haskell processes with full-fledged
parallelism.

Hence, the performance measurements in Figs.~\ref{fig:bench-rm-sm}--\ref{fig:sudokuSMBenchmark} and table~\ref{tab:bestAndWorstResults} use only real cores as the purpose of this paper is to show the performance of PArrows and not to investigate parallel performance with hyper-threading in use.
